# Scatterplots, boxplots, and violin-boxplots {#07-more-viz}

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE)

# Load the tidyverse package below
library(tidyverse)

# Load the data file
# This should be the Zhang_2014.csv file 
zhang_data <- read_csv("data/Zhang_2014.csv")

zhang_data <- zhang_data %>%
  select(Gender, Age,Condition, T1_Predicted_Interest_Composite, T2_Actual_Interest_Composite) %>%
  mutate(subject = row_number()) %>%
  pivot_longer(names_to = "time",values_to = "interest",
               cols = T1_Predicted_Interest_Composite:T2_Actual_Interest_Composite) %>%
  mutate(Condition = recode(Condition, 
                            "1" = "Ordinary", 
                            "2" = "Extraordinary")) %>%
  mutate(time = recode(time, 
                       "T1_Predicted_Interest_Composite" = "time1_interest", 
                       "T2_Actual_Interest_Composite" = "time2_interest"),
         Gender = recode(Gender, 
                         "1" = "male", 
                         "2" = "female")) %>%
  filter(Gender %in% c("male", "female"))
```

In this chapter we are going to focus on visualising data using `ggplot2`. You've already created a number of different plots including bar charts, scatterplots, histograms, qq-plots, and violin-boxplots, but now we will show you how to customise your plots further to give you a better idea of the range and flexibility of visualising data in R. 

**Chapter Intended Learning Outcomes (ILOs)**

By the end of this chapter, you will be able to: 

- ILO1.

## Chapter preparation

### Introduction to the data set 

For this chapter, we are using open data from @zhang_present_2014. The abstract of their article is:

> Although documenting everyday activities may seem trivial, four studies reveal that creating records of the present generates unexpected benefits by allowing future rediscoveries. In Study 1, we used a time-capsule paradigm to show that individuals underestimate the extent to which rediscovering experiences from the past will be curiosity provoking and interesting in the future. In Studies 2 and 3, we found that people are particularly likely to underestimate the pleasure of rediscovering ordinary, mundane experiences, as opposed to extraordinary experiences. Finally, Study 4 demonstrates that underestimating the pleasure of rediscovery leads to time-inconsistent choices: Individuals forgo opportunities to document the present but then prefer rediscovering those moments in the future to engaging in an alternative fun activity. Underestimating the value of rediscovery is linked to people’s erroneous faith in their memory of everyday events. By documenting the present, people provide themselves with the opportunity to rediscover mundane moments that may otherwise have been forgotten.

In summary...

### Organising your files and project for the chapter

Before we can get started, you need to organise your files and project for the chapter, so your working directory is in order.

1. In your folder for research methods and the book `ResearchMethods1_2/Quant_Fundamentals`, create a new folder called `Chapter_07_dataviz`. Within `Chapter_07_dataviz`, create two new folders called `data` and `figures`.

2. Create an R Project for `Chapter_07_dataviz` as an existing directory for your chapter folder. This should now be your working directory.

3. Create a new R Markdown document and give it a sensible title describing the chapter, such as `07 Scatterplots Boxplots Violins`. Delete everything below line 10 so you have a blank file to work with and save the file in your `Chapter_07_dataviz` folder. 

4. We are working with a new data set, so please save the following data file: [Zhang_2014.csv](data/Zhang_2014.csv). Right click the link and select "save link as", or clicking the link will save the files to your Downloads. Make sure that you save the file as ".csv". Save or copy the file to your `data/` folder within `Chapter_07_dataviz`.

You are now ready to start working on the chapter! 

### Activity 1 - Read and wrangle the data

As the first activity, try and test yourself by completing the following task list to practice your data wrangling skills. If you just want to focus on data visualisation, then you can just type the code in the solution. 

```{r eval=FALSE}
# Load the tidyverse package below
?
  
# Load the two data files
# This should be the ahi-cesd.csv file 
dat <- ?

# This should be the participant-info.csv file
pinfo <- ?
```

::: {.callout-tip collapse="true"}
#### Show me the solution
You should have the following in a code chunk: 

```{r eval=FALSE}
# Load the tidyverse package below
library(tidyverse)

# Load the data file
# This should be the Zhang_2014.csv file 
zhang_data <- read_csv("data/Zhang_2014.csv")

zhang_data <- zhang_data %>%
  select(Gender, Age,Condition, T1_Predicted_Interest_Composite, T2_Actual_Interest_Composite) %>%
  mutate(subject = row_number()) %>%
  pivot_longer(names_to = "time",values_to = "interest",
               cols = T1_Predicted_Interest_Composite:T2_Actual_Interest_Composite) %>%
  mutate(Condition = recode(Condition, 
                            "1" = "Ordinary", 
                            "2" = "Extraordinary")) %>%
  mutate(time = recode(time, 
                       "T1_Predicted_Interest_Composite" = "time1_interest", 
                       "T2_Actual_Interest_Composite" = "time2_interest"),
         Gender = recode(Gender, 
                         "1" = "male", 
                         "2" = "female")) %>%
  filter(Gender %in% c("male", "female"))
```

:::

## Activity 3: Scatterplots {#viz-a3}

### Basic scatterplot

Now let's make a scatterplot plotting `Age` and `interest` to see if there is any relationship between the two. We need to specify both the x and y-axis variables. The following code will produce a very simple scatterplot. Again, you wouldn't use this graph in a paper, but for eye-balling your data it would suffice. 

```{r scat1, fig.cap = "Basic scatterplot"}
ggplot(zhang_data, aes(x = interest,y = Age))+
       geom_point()
```

### Axis labels

From this plot it doesn't look like there is much of a relationship between age and interest ratings. We can now change the labels using the same scale functions as before. 

```{r scat2, fig.cap = "Scatterplot with label changes"}
ggplot(zhang_data, aes(x = interest,y = Age))+
       geom_point()+
  scale_x_continuous(name = "Mean interest score (1-7)") + 
  scale_y_continuous(name = "Age")
```

### Adding a regression line

It's often useful to add a regression line or line of best fit to a scatterplot. The regression line is added with `geom_smooth()` and by default will also provide a 95% confidence interval. You can specify what type of line you want to draw, most often you will need `method = lm`, i.e., a linear model or a straight line. Look up the help documentation for `geom_smooth()` and see what other methods you can use. 

```{r scat3, fig.cap = "Scatterplot with regression line"}
ggplot(zhang_data, aes(x = interest,y = Age))+
  geom_point()+
  scale_x_continuous(name = "Mean interest score (1-7)") + 
  scale_y_continuous(name = "Age")+
  geom_smooth(method=lm) # if you don't want the shaded CI, add se = FALSE to this
```

### Grouped scatterplots

We can use ggplot to show how the relationship might differ for different populations within our data. We do this by adding `colour()` to `aes()` and setting it as whatever variable we would like to distinguish between. In this case, we will see how the relationship between age and interest differs for the male and female participants. There are a few participants with missing gender so we will first filter them out.


```{r, scat4, fig.cap = "Grouped scatterplot", warning=FALSE}
zhang_data %>%
  filter(Gender %in% c("male", "female")) %>%
           ggplot(aes(x = interest,y = Age, colour = Gender))+
  geom_point()+
  scale_x_continuous(name = "Mean interest score (1-7)") + 
  scale_y_continuous(name = "Age")+
  geom_smooth(method=lm)
```

And here's that plot with the labels tidied up. Notice the use of `scale_color_discrete()` to adjust the labels for Gender. 

```{block, type="warning"}
When you change the `labels`, R will simply overwrite the names in the dataset. If you wanted to actually change the order of the categories (e.g., have male as the red line) you need to change the order of the factor. We will do this later, for now, just be sure that you're changing the name of the right category (i.e., female comes first))
```

```{r scat5, fig.cap = "Grouped scatterplot with adjusted labels"}
ggplot(zhang_data, aes(x = interest,y = Age, colour = Gender))+
  geom_point()+
  scale_x_continuous(name = "Mean interest score (1-7)") + 
  scale_y_continuous(name = "Age")+
  geom_smooth(method=lm)+
  scale_color_discrete(name = "Gender",
                       labels = c("Female", "Male"))
```

## Activity 4: Boxplots {#viz-a4}

### Basic boxplot

The following code will produce a simple boxplot for eye-balling your data.

```{r bp1, fig.cap = "Basic boxplot"}
ggplot(zhang_data, aes(x = Condition, y = interest))+
  geom_boxplot()
```

### Adding data points

If we add another layer `geom_point()` we can add our raw data points to our boxplots to make them more informative.

```{r bp2, fig.cap = "Boxplot with overplotting"}
ggplot(zhang_data, aes(x = Condition, y = interest))+
  geom_boxplot()+
  geom_point()
```

However, this plot suffers from **over-plotting**, that is, there are multiple data points on top of each other. We can change this by using `geom_jitter()`, which adds a layer of points that are jittered so that each one is visible.

`height` and `width` affect how much each point is jittered. Play around with the values to see how it affects the data points. 

```{r bp3, fig.cap = "Boxplot with jittered data"}
ggplot(zhang_data, aes(x = Condition, y = interest))+
  geom_boxplot()+
  geom_jitter(height = 0, width = .1)
```

### Adding colour

We may want to add colour to our graph (and for consistency, we'll sort out the labels). We do this by adding the 'fill' argument to the ggplot aesthetic by specifying which variable the colour of the fill should be organised by. 

```{r bp4, fig.cap = "Boxplot with colour"}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = Condition))+
  geom_boxplot()+
  geom_jitter(height = 0, width = .1)+
  scale_x_discrete(name = "Condition") + # note the x-axis is discrete
  scale_y_continuous(name = "Mean interest rating (1-7)")+
  scale_fill_discrete(guide = FALSE) # this suppresses the legend because we don't need it
```

### Boxplots for multiple factors

When you only have one IV, using the fill command to change the colour is a little redundant, as the colours don't add any additional information. It makes more sense to use colour to represent an additional IV. 

For this example, we'll use `Condition` and `time` as IVs. `fill()` now specifies a second IV, rather than repeating the IV on the x-axis as in the previous plot. 

With multiple IVs the command to overlay the raw data points changes as the data points also need dodged (try running the code with the previous geom_jitter function to see what happens)

```{r bp5, fig.cap = "Boxplot for two factors"}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = time))+
  geom_boxplot()+
  geom_point(position=position_jitterdodge(jitter.width = .1))
```

### Colour-blind friendly options

There is one more fill option that we can use. Rather than specifying `scale_fill_discrete()`, we can use `scale_fill_viridis_d()`. This function does exactly the same thing but it uses a colour-blind friendly palette (which also prints in black and white). There are 5 different options for colours and you can see them by changing `option` to A, B, C, D or E. Personally I like option E with `alpha = .6` (to control transparency) but that's not an official School position.

```{r, fig.cap= "Boxplots with friendly colours"}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = time))+
  geom_boxplot(alpha = .6)+
  geom_point(position=position_jitterdodge(jitter.width = .1)) +
  scale_fill_viridis_d(option = "E")
```

## Activity 5: Reordering factors {#viz-a5}

R orders categorical variables alphabetically. For gender it didn't really matter whether male or female was represented first and for time 1 and 2 it makes sense for them to be in this order but we may want to change the order of Condition (in my mind it makes more sense for Ordinary to come first, but that may just be me).

To do this we can use `mutate()` and `fct_level()` to change the factor levels to the order we want.

```{r reorder}
zhang_data <- zhang_data %>%
  mutate(Condition = fct_relevel(Condition, c("Ordinary", "Extraordinary")))
```

Now we can re-run the boxplot. That's better. 

```{r, bp6, fig.cap = "Boxplot with reordered factors"}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = time))+
  geom_boxplot(alpha = .6)+
  geom_point(position=position_jitterdodge(jitter.width = .1)) +
  scale_fill_viridis_d(option = "E")
```

## Activity 7: Violin plots {#viz-a7}

### Basic violin plot

Violin plots are so-called because with a normal distribution the shape would look something like a violin. They show density, i.e., the fatter the violin the more data points there are for that value.

```{r vp1, fig.cap = "Basic violin plot"}
ggplot(zhang_data, aes(x = Condition, y = interest))+
  geom_violin()
```

### Violin plots with raw data points

Like the boxplot, we can also add the raw data points to our violin plot, making sure to use jitter  to avoid over-plotting.

```{r vp2, fig.cap = "Violin plot with data points"}
ggplot(zhang_data, aes(x = Condition, y = interest))+
  geom_violin()+
  geom_jitter(height = 0, width = .1)
```

```{block, type="warning"}
It's important to remember that R is very literal. `ggplot2` works on a system of layers. It will add new geoms on top of existing ones and it won't stop to think whether this is a good idea. Try running the above code but put `geom_jitter()` first and then add `geom_violin()`. The order of your layers matters.
```

## Activity 8: Violin-boxplots {#viz-a8}

One increasingly common graph is a violin + boxplot + summary plot that shows a huge amount of information about your data in a single plot. 

* This code uses two calls to `stat_summary()` that was introduced during the t-test chapter. The first draws a `point` to represent the mean, and the second draws an `errorbar` that represents standard error (`mean_se`).  
* `guides` is a new function and can be used to adjust whether legends are displayed. This has the same effect as specifying `show.legend = FALSE` in both `geom_violin()` and `geom_boxplot()` but it uses less code to do so.  
* `fatten = NULL` removes the median line from the boxplots. This can be useful if you're running a test where you're comparing means as it makes it easier to see the point range. 
* You may get warning messages telling you that R has removed rows containing missing values, you do not need to worry about this.

```{r vbp1, fig.cap = "Violin-boxplot with summary data"}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = Condition))+
  geom_violin(alpha = .6, trim = FALSE)+
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) +
  scale_fill_viridis_d(option = "E", label = c("Ordinary", "Extraordinary"))+
  scale_y_continuous(name = "Mean interest rating (1-7)") +
  guides(fill = FALSE)

```

## Activity 9: Faceting {#viz-a9}

`ggplot2` contains a facet function that produces different plots for each level of a grouping variable which can be very useful when you have more than two factors, for example, for a three-way ANOVA. The following code displays produces violin-boxplots for Condition ~ interest, but separately for male and female participants. 

* This code adds an extra argument `position = position_dodge(.9)` to align the layers with the violin plots. Try removing this argument from each layer to see what happens, and also try adjusting the value from `.9` to another number.

```{r facet1, fig.cap = "Violin-boxplot facetted by gender", message = FALSE}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = time))+
  geom_violin(alpha = .6, trim = FALSE)+
  geom_boxplot(width = .2, 
               alpha = .6, 
               fatten = NULL,
               position = position_dodge(.9))+
  stat_summary(fun = "mean", geom = "point",
               position = position_dodge(.9)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1,
               position = position_dodge(.9))+
  scale_fill_viridis_d(option = "E") +
  facet_wrap(~Gender)
```

### Facet labelling

Finally, changing the labels within the facets is a little more complicated - there's no additional scale layer, instead, you adjust this inside `facet_wrap()` using `labeller`. This has always felt unintuitive to me and I have to look it up every single time so don't worry if it is confusing - just remember where to look for the example. 

```{r facet2, fig.cap = "Facetted plot with updated labels"}
ggplot(zhang_data, aes(x = Condition, y = interest, fill = time))+
  geom_violin(alpha = .6, trim = FALSE)+
  geom_boxplot(width = .2, 
               alpha = .6, 
               fatten = NULL,
               position = position_dodge(.9))+
  stat_summary(fun = "mean", geom = "point",
               position = position_dodge(.9)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1,
               position = position_dodge(.9))+
  scale_fill_viridis_d(option = "E") +
  facet_wrap(~Gender, labeller = labeller(Gender = (c(female = "Female", male = "Male"))))
```

## Test yourself

To end the chapter, we have some knowledge check questions to test your understanding of the concepts we covered in the chapter. We then have some error mode tasks to see if you can find the solution to some common errors in the concepts we covered in this chapter. 

### Knowledge check

**Question 1**. We have...

`r longmcq(sample(c("select()", answer="filter()", "mutate()", "arrange()", "group_by()", "summarise()")))`

**Question 2**. We are...

`r longmcq(sample(c("select()", "filter()", "mutate()", "arrange()", "group_by()", answer="summarise()")))`

**Question 3**. Our data..

`r longmcq(sample(c("select()", "filter()", answer="mutate()", "arrange()", "group_by()", "summarise()")))`

**Question 4**. We want...

`r longmcq(sample(c(answer="group_by() and summarise()", "filter() and select()", "group_by() and arrange()", "arrange() and mutate()")))` 

**Question 5**. If we...

`r longmcq(sample(c(answer = "wide_data %>% select() %>% pivot_longer() %>% group_by() %>% summarise()", "long_data %>% select() %>% pivot__longer_wider() %>% group_by() %>% summarise()", "wide_data %>% pivot_longer() %>% select() %>% summarise() %>% group_by()", "select() %>% pivot_longer() %>% group_by() %>% summarise() %>% wide_data")))`

### Error mode

The following questions are designed to introduce you to making and fixing errors. For this topic, we focus on the new types of data visualisation. Remember to keep a note of what kind of error messages you receive and how you fixed them, so you have a bank of solutions when you tackle errors independently. 

Create and save a new R Markdown file for these activities. Delete the example code, so your file is blank from line 10. Create a new code chunk to load <pkg>tidyverse</pkg> and the data files: 

```{r eval=FALSE}
# Load the tidyverse package below
library(tidyverse)

# Load the data file
# This should be the Zhang_2014.csv file 
zhang_data <- read_csv("data/Zhang_2014.csv")
```

Below, we have several variations of a code chunk error or misspecification. Copy and paste them into your R Markdown file below the code chunk to load <pkg>tidyverse</pkg> and the data files. Once you have copied the activities, click knit and look at the error message you receive. See if you can fix the error and get it working before checking the answer.

**Question 6**. Copy the following code chunk into your R Markdown file and press knit. 

````{verbatim, lang = "markdown"}
```{r}

```
````

::: {.callout-caution collapse="true"} 
#### Explain the solution


```{r eval = FALSE}

```
:::

**Question 7**. Copy the following code chunk into your R Markdown file and press knit. 

````{verbatim, lang = "markdown"}
```{r}

```
````

::: {.callout-caution collapse="true"} 
#### Explain the solution


```{r eval = FALSE}

```
:::

**Question 8**. Copy the following code chunk into your R Markdown file and press knit.

````{verbatim, lang = "markdown"}
```{r}

```
````

::: {.callout-caution collapse="true"} 
#### Explain the solution

```{r eval = FALSE}

```
:::

## Words from this Chapter

Below you will find a list of words that were used in this chapter that might be new to you in case it helps to have somewhere to refer back to what they mean. The links in this table take you to the entry for the words in the [PsyTeachR Glossary](https://psyteachr.github.io/glossary/){target="_blank"}. Note that the Glossary is written by numerous members of the team and as such may use slightly different terminology from that shown in the chapter.

```{r gloss, echo=FALSE, results='asis'}
glossary_table()
```

## End of chapter

Blah. 

